Apache Flink\texttrademark: Unified Stream and Batch Processing \\ in a Single Engine

### 2015-11-29 batch.tex
Small changes
remake iteration sections

\subsection{Batch Iterations on top of Dataflows}
\label{sec:batch-iterations}
Most batch dataflow approaches execute iterations by submitting a new job for each iteration or by adding additional nodes to a running DAG \cite{DBLP:journals/pvldb/BuHBE10, DBLP:conf/hotcloud/ZahariaCFSS10}. Flink keeps the DAG-based runtime and scheduler, and allows for special ``head'' and ``tail'' tasks that are connected with feedback edges, to signify the beginning and end of each iteration. Flink's iterations are competitive with specialized graph engines \cite{DBLP:journals/pvldb/EwenTKM12}. This is achieved through the concept of  \emph{delta} iterations, which exploit sparse computational dependencies, and are used as the basis for Gelly, Flink's Graph API. Finally, Flink offers Bulk Synchronous Parallel (BSP) as well as Stale Synchronous Parallel (SSP) iterations in its DataSet API.

>>>>

\subsection{Batch Iterations on top of Dataflows}
\label{sec:batch-iterations}
Structured iteration support is the basis of most distributed batch processing applications. Iterative graph analytics, parallel gradient descent and optimisation techniques have been implemented in the past on top of Bulk Synchronous Parallel (BSP) and Stale Synchronous Parallel (SSP) models, among others. The Flink execution model, allows for any type of structured iteration logic using iteration control events, as explained in \ref{sec:iterations}. For example in the case of a BSP execution, iteration control events mark the beginning and the end of supersteps in an iterative computation. Finally, Flink introduces further novel optimisation techniques such as the concept of \emph{delta} iterations \cite{DBLP:journals/pvldb/EwenTKM12}, which can exploit sparse computational dependencies (already part of Gelly, Flink's Graph API).

### 2015-11-29 exec-model.tex
fix figure and add some example references
remake iteration sections

An operator receives barriers from upstream and first performs an alignment phase, making sure that the barriers from all inputs have been received. Then, the operator writes its state (e.g., contents of a sliding window, or custom data structures) to durable storage (the storage backend can be an external system, e.g., HDFS). Once the state has been backed up, the operator forwards the barrier downstream. ABS bears resemblances to the seminal Chandy-Lamport algorithm for asynchronous distributed snapshots \cite{chandy1985distributed}. However, because of the DAG structure of a Flink program, ABS does not need to checkpoint in-flight records but solely rely on the aligning phase to apply all their effects to the operator states. This guarantees that the data that needs to be written to reliable storage is kept to the theoretical minimum (i.e., only the current state of the operators).

@@

>>>>

An operator receives barriers from upstream and first performs an alignment phase, making sure that the barriers from all inputs have been received. Then, the operator writes its state (e.g., contents of a sliding window, or custom data structures) to durable storage (the storage backend can be an external system, e.g., HDFS). Once the state has been backed up, the operator forwards the barrier downstream. Eventually, all operators will register a snapshot of their state and a global snapshot will be complete. For example, in \autoref{fig:snapshots} we show that snapshot $t2$ contains all operator states that are the result of consuming all records before \emph{t2 barrier}. ABS bears resemblances to the seminal Chandy-Lamport algorithm for asynchronous distributed snapshots \cite{chandy1985distributed}. However, because of the DAG structure of a Flink program, ABS does not need to checkpoint in-flight records but solely rely on the aligning phase to apply all their effects to the operator states. This guarantees that the data that needs to be written to reliable storage is kept to the theoretical minimum (i.e., only the current state of the operators).

@@

\subsection{Iterative Dataflows}
\label{sec:iterations}

\begin{figure}[t!]
  \centering
    \includegraphics[width=.75\textwidth]{figs/iterations.pdf}
    \vspace{-6mm}
  \caption{The iteration model of Apache Flink}
  \vspace{-2mm}
  \label{fig:iterations}
\end{figure}

Loops and iterations are crucial for several applications. Support for iterations in data-intensive computing systems typically relies on submitting a new job for each iteration or by adding additional nodes to a running DAG \cite{DBLP:journals/pvldb/BuHBE10, DBLP:conf/hotcloud/ZahariaCFSS10}. As depicted in \autoref{fig:iterations}, Flink maintains the DAG-based runtime and scheduler, and allows for special iteration ``head'' and ``tail'' tasks that are implicitly connected with feedback edges. The role of these tasks is to establish an active feedback channel to the DAG and provide coordination for processing data records in transit within this feedback channel. Coordination is needed for implementing any type of structured parallel iteration model such as the Bulk Synchronous Parallel (BSP) model as we explain in section \ref{sec:batch-iterations}.

### 2015-11-29 streaming.tex
remake iteration sections

\subsection{Asynchronous Streaming Iterations}
Stream loops are crucial for several applications. Common examples are feedback loops for incrementally building and training machine learning models, reinforcement learning, and executing graph algorithms. Flink supports loops through explicit feedback edges in its dataflow graph, allowing the implementation of asynchronous iterative/incremental algorihtms. In Flink, feedback streams have special semantics and are treated in a way that is transparent to its scheduler: feedback streams do not act as dependencies in the dataflow graph, allowing Flink to employ simple DAG scheduling techniques while the contents of feedback streams are treated as operator state, in order to simplify fault tolerance \cite{carbone2015lightweight}. Flink allows constructing loops on top of its DataStream API, and can trivially subsume structured loops such as differential dataflows \cite{murray2013naiad}. \autoref{sec:batch-iterations} describes iterations support for batch analytics.

>>>>

\subsection{Asynchronous Streaming Iterations}
Stream loops are essential for several applications. Common examples are feedback loops for incrementally building and training machine learning models, reinforcement learning and graph algorithms in the semi-streaming model \cite{feigenbaum2005graph}.
In most such cases, feedback loops need no coordination. Asynchronous iterations cover the communication needs for streaming applications and differ from parallel optimisation problems that are based on structured iterations on finite data. As presented, in \ref{sec:iterations} and \ref{fig:iterations} the execution model of Apache Flink already covers asynchronous iterations, when no iteration control mechanism is enabled. In addition, to comply with fault-tolerance guarantees feedback streams are treated as operator state within the implicit iteration head operator and are part of a global snapshot, as explained in detail in \cite{carbone2015lightweight}. The DataStream API allows for an explicit definition of feedback streams and can trivially subsume future support for structured loops over streams such as differential dataflows \cite{murray2013naiad}.
